Demography and selection since maize domestication

### 2015-05-15 manuscript.tex
updates

The most likely model suggested an ancestral for $\theta$ ($4N\mu$) of
0.014734, which is similar to previous estimates
\textcolor{red}{(references here)}. Assuming a mutation rate of $\mu =
3 \times 10^{-8}$ \textcolor{red}{(reference here)}, this suggests an effective
population size of $N_a = 122,783$ individuals, a population split
$T_B = 15,523$ generations in the past, a maize bottleneck
effective populations size of $N_b = 6,455$ individuals (5.26\% of
$N_a$), a modern maize effective population size of $N_m = 366,973$
individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
maize, and $M_{mt} = 1.72$ migrants per generation from maize to
teosinte (Figure \ref{bottleneck}). We note that a population
split over 15 thousand generations before present precedes estimates
from
archaeological data which suggests maize domestication began
approximately 9,000 years before present \textcolor{red}{(reference
  here)}. This could result from multiple generations per year, or in
may reflect teosinte population structure that was present before
domestication. Also, note that the genetic time of the population
split must precede morphological changes that could be identified
morphologically.   Additionally, because recent expansion is
most evidenced by rare alleles, and because these data provide low
power to detect rare alleles, we expect that the estimate of  $N_m =
366,973$, or $\sim 3N_a$, is likely an underestimate. \textcolor{red}{(This
  is where evaluation using Kate's SFS should go)}.

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{This is what domestication looked like. WRITE CAPTION!}
\label{bottleneck}
\end{figure}

>>>>

The most likely model suggested an ancestral for $\theta$ ($4N\mu$) of
0.014734, which is similar to previous estimates
\textcolor{red}{(references here)}. Assuming a mutation rate of $\mu =
3 \times 10^{-8}$ \textcolor{red}{(reference here)}, this suggests an effective
population size of $N_a = 122,783$ individuals, a population split
$T_B = 15,523$ generations in the past, a maize bottleneck
effective populations size of $N_b = 6,455$ individuals (5.26\% of
$N_a$), a modern maize effective population size of $N_m = 366,973$
individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
maize, and $M_{mt} = 1.72$ migrants per generation from maize to
teosinte (Figure \ref{bottleneck}). We note that a population
split over 15 thousand generations before present precedes estimates
from
archaeological data which suggests maize domestication began
approximately 9,000 years before present \textcolor{red}{(reference
  here)}. This could result from multiple generations per year, or in
may reflect teosinte population structure that was present before
domestication. Also, note that the genetic time of the population
split must precede morphological changes that could be identified
morphologically.   Additionally, because recent expansion is
most evidenced by rare alleles, and since these data provide low
power to detect rare alleles, we expect that the estimate of  $N_m =
366,973$, or $\sim 3N_a$, is likely an underestimate.

We utilized a complementary dataset of 4,808 maize landrace samples collected
from across the Americas \textcolor{red}{(SeeDs reference here)} to
obtain a less downward-biased estimate of the current maize effective
population size. According to a singleton-based estimate of $\theta$
\cite{fu1993}, which was chosen since rapid post-bottleneck expansion
implies that rare alleles will most accurately reflect current maize
demography, we obtained the estimate \textcolor{red}{$N_e =
  1,011,686$}. Although less biased than the previously mentioned
esitmate, we note that ascertainment bias of the
genotyping platform used to generate these data again biases this
estimate downward.

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{Parameters of domestication as estimated by dadi. Notably,
  the maize effective population size ($N_e$) during the domestication
  bottleneck appears to have consisted of approximately 5.26\%
  the ancestral $N_e$, before recovering two at least 2.98
  times as large as the ancestral $N_e$.}
\label{bottleneck}
\end{figure}
